Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-11-29 dsn-2018/bare_conf.tex
Expand Introduction, Related Work...
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

Cryptographic hash functions form the core of many protocols. From file
integrity checks to verify long term storage of data, to cache invalidation
techniques, and use as a building block in network protocols such as Kerberos
and TLS, this class of functions necessarily has some strong guarantees about
the properties of these functions. There are three basic properties hash
functions must have to be considered cryptographically secure:
\begin{itemize}
    \item Preimage Resistance: It should be computationally hard to find
        the inverse of a hash function.
    \item Second Preimage Resistance: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item Collision Resistance: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that a second preimage is necessarily a collision and is also a stronger
result.

@@

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use. In some
instances however, there continues to be widespread use of deprecated hash
functions. One widespread example is the continued use by Git of SHA-1, despite
M. Stevens et al. having published the first known full SHA-1 collision in
February of 2017 \cite{cryptoeprint:2017:190}. However, this attack had a bound
of $2^{63.1}$ -- faster than the theoretical $2^{80}$ but still requiring
significant investment in compute resources.

\textbf{The contributions of this paper are...}

\begin{itemize}
\item We...
\end{itemize}

We define a series of criteria for analyzing the utility of a collision class.
Collision classes of high utility are closer to a second preimage attack in the
amount of flexibility they provide to an attacker, whereas classes of low
utility may not affect all systems which use a particular hash function.
Furthermore, we demonstrate new techniques for working with collisions under
a logical cryptanalysis framework. Throughout this paper, we make use of the
MD4 hash function due to the ease of generating new classes of collisions.

The remainder of this paper is organized as follows... In Section
\ref{Sec:Terminology} we...

@@

Broad context... intersection... subsection introductions.

@@

\subsection{Satisfiability Solving}

@@

\section{Measuring the Utility of a Collision}

@@

\section{Techniques for Logical Cryptanalysis}

@@

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric.

@@

\begin{table}
    \caption{Distance between Existing Collision Classes}
    \label{table:distance}
    \begin{tabular}{c c c c c c}
        & X. W. & Y. S. & P. K. & H. D. & M. S. \\
        X. Wang's & 0 & 21 & 26 & 27 & 12 \\
        Y. Sasaki's & & 0 & 25 & 26 & 2 \\
        P. Kasselman's &  &  & 0 & 1 & 25 \\
        H. Dobbertin's &  &  &  & 0 & 26 \\
        M. Schlaffer's &  &  &  &  & 0 \\
    \end{tabular}
\end{table}

@@

\section{New Collision}

\section{Discussion}

@@

\section{Conclusion}

@@

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank the Department of Electrical and Computer
Engineering and Nathan Mulbrook for providing hardware.

>>>>

Cryptographic hash functions form the core of many protocols. From file
integrity checks to verify long term storage of data, to cache invalidation
techniques, and use as a building block in network protocols such as Kerberos
and TLS, this class of functions necessarily has strong guarantees about
properties of its members. There are three basic properties hash
functions must have to be considered cryptographically secure:
\begin{itemize}
    \item \textbf{Preimage Resistance}: It should be computationally hard to find
        the inverse of a hash function.
    \item \textbf{Second Preimage Resistance}: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item \textbf{Collision Resistance}: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that a second preimage is necessarily a collision and is also a stronger
result in practice.

@@

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use. In some
instances however, there continues to be widespread use of deprecated hash
functions. One widespread example is the continued use by Git of SHA-1, despite
M. Stevens et al. publishing a full SHA-1 collision in February of 2017
\cite{cryptoeprint:2017:190}. However, this attack had a bound of $2^{63.1}$ --
faster than the theoretical $2^{80}$ but still requiring significant compute
resources.

We seek to address the problem of evaluating the use cases for a collision,
to determine--or partially determine--whether or not a collision affects
a particular system. This paper contributes several novel techniques to
logical cryptanalysis and uses these techniques to derive metrics of the
trustworthiness of hash functions. We claim the following results are novel
in the field:
\begin{itemize}
    \item We define the neighborhood of a class of collisions and show that
        it is frequently non-empty.
    \item We discuss techniques for finding useful collisions within a
        class of collisions.
    \item We propose additional techniques for building new classes of
        collisions and show that collisions in MD4 can be found inductively.
    \item We propose five universal metrics for evaluating the utility of
        a class of collisions.
\end{itemize}

\textbf{TODO - Transition}

We propose that collision classes of high utility are closer to a second
preimage attack in the amount of flexibility they provide to an attacker,
whereas classes of low utility may not affect all systems which use a
particular hash function. Furthermore, we demonstrate new techniques for
working with collisions under a logical cryptanalysis framework. Throughout
this paper, we make use of the MD4 hash function due to the ease of generating
new classes of collisions.

\textbf{TODO - Finish Organization}

The remainder of this paper is organized as follows. In Section
\ref{Sec:Related} we discuss previous results and show how our research
expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
terminology and notations we use throughout the remainder of the paper.
In Section \ref{Sec:Intuition}, we give the intuition between measuring
the utility of a collision and motivate why it is useful. In Section
\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
analyze collisions in hash functions.

@@

\textbf{Broad context... intersection... subsection introductions.}

This paper draws upon the public cryptanalysis of MD4, the wide availability of
high quality SAT solvers, and previous work in logical cryptanalysis.

@@

Differential Cryptanalysis has been the major technique behind the discovery
of collisions in cryptographic hash functions. Its importance can be seen
everywhere from X. Wang's attacks on MD4 \cite{Wang2005} to more recent
cryptanalysis of general purpose hash functions such as Murmur3 by
J. Aumasson, D. Bernstein, and M. Bo{\ss}let \cite{murmur3DC}. However, while
differential cryptanalysis plays an important role, techniques for automated
analysis of hash functions are still an emerging area. In
\cite{cryptoeprint:2004:199}, X. Wang claimed to have found the MD4 collision
``with hand calculation'', though in \cite{Wang2005}, she specifies message
modification techniques which can be implemented in code.

Our work removes the need for finding message modification techniques by
specifying the differential path as part of the SAT formula and letting the
solver find a pair of satisfying messages which follow the differential path
and produce a collision. Further constraints can then be placed upon this
model, such as a chosen prefix or desired start state. We believe performing
these attacks by hand to be difficult, and note that message modification
techniques may affect the chosen prefix.

\subsection{Logical Cryptanalysis}

Logical Cryptanalysis and its encoding as SAT likely started under the work of
F. Massacci in early 2000 with his paper titled ``Logical Cryptanalysis as a
SAT Problem: Encoding and Analysis of the U.S. Data Encryption Standard''.

\subsection{Trustworthy Computing}

\textbf{TODO - Assigned @erozier}

@@

\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}

@@

\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}

@@

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric. We extend this distance
function to include an \textit{absolute} distance, measuring the number of
intermediate rounds with non-zero differences. We denote this as $\delta(C_1)$.

@@

\begin{table}
    \caption{Distance between Existing Collision Classes}
    \label{table:distance}
    \begin{tabular}{c c c c c c c}
        & X. W. & Y. S. & P. K. & H. D. & M. S. & Absolute \\
        X. Wang's & 0 & 21 & 26 & 27 & 12 & 18 \\
        Y. Sasaki's & & 0 & 25 & 26 & 2 & 16 \\
        P. Kasselman's &  &  & 0 & 1 & 25 & 11 \\
        H. Dobbertin's &  &  &  & 0 & 26 & 12 \\
        M. Schlaffer's &  &  &  &  & 0 & 17 \\
    \end{tabular}
\end{table}

@@

\section{New Collision} \label{Sec:Collision}

\section{Discussion of Impacts} \label{Sec:Impacts}

@@

\section{Conclusion} \label{Sec:Conclusion}

@@

% use section* for acknowledgment
\section*{Acknowledgments}

The authors would like to thank the Department of Electrical and Computer
Engineering and XXXXX XXXXX for providing hardware.
